PLAE web app enables powerful searching and multiple visualizations across one million unified single-cell ocular transcriptomes

### 2022-09-12 manuscript_00.Rmd
update

date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  word_document:
    reference_docx: word-styles-reference-01.docx
    fig_caption: yes
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
bibliography: zotero_plaeApp.bib
csl: investigative-ophthalmology-and-visual-science.csl
abstract: "PURPOSE: To create a high performance reactive web application to query single cell gene expression data across cell type, species, study, and other factors. METHODS: We updated the content and structure of the underlying data (single cell Eye in a Disk, scEiaD) and wrote the web application PLAE (https://plae.nei.nih.gov) to visualize and explore it. RESULTS: The new portal provides quick visualization of over a million individual cells from vertebrate eye-and body transcriptomes encomopassing four species, 60 cell types, six ocular tissues, and 23 body tissues across 37 publications. To demonstrate the value of this unified pan-eye dataset, we replicate known neurogenic and cone macula markers as well as propose six???? new cone human region markers. CONCLUSION: The PLAE web application provides the eye community a powerful and quick means to test hypotheses related to gene expression across a highly diverse, community-derived database."

@@

## Data reproducibility

The code base for scEiaD is available at github.com/davemcg/scEiaD; the commit corresponding to this manuscript is 8f2c04b. There are three relevant Snakemake pipelines which were used to create scEiaD: SnakeQUANT, which was used to quantify the gene / cell expression from the raw fastq files, SnakePOP, which runs the multi-parameter integration methods, and SnakeSCEIAD, which uses the optimal parameters identified in SnakePOP to run the cell type machine learning and differential gene testing that is incorporated into the scEiaD sqlite database for the app. The code base for the web app (along with local installation instructions) is available at github.com/davemcg/scEiaD/plaeApp.

>>>>

date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  word_document:
    reference_docx: word-styles-reference-01.docx
    fig_caption: yes
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
bibliography: zotero_plaeApp.bib
csl: investigative-ophthalmology-and-visual-science.csl
abstract: "PURPOSE: To create a high performance reactive web application to query single cell gene expression data across cell type, species, study, and other factors. METHODS: We updated the content and structure of the underlying data (single cell Eye in a Disk, scEiaD) and wrote the web application PLAE (https://plae.nei.nih.gov) to visualize and explore it. RESULTS: The new portal provides quick visualization of over a million individual cells from vertebrate eye-and body transcriptomes encomopassing four species, 60 cell types, six ocular tissues, and 23 body tissues across 37 publications. To demonstrate the value of this unified pan-eye dataset, we replicate known neurogenic and cone macula markers as well as propose six new cone human region markers. CONCLUSION: The PLAE web application provides the eye community a powerful and quick means to test hypotheses related to gene expression across a highly diverse, community-derived database."

@@

## Data reproducibility and availability

The code base for scEiaD is available at github.com/davemcg/scEiaD; the commit corresponding to this manuscript is 8f2c04b. There are three relevant Snakemake pipelines which were used to create scEiaD: SnakeQUANT, which was used to quantify the gene / cell expression from the raw fastq files, SnakePOP, which runs the multi-parameter integration methods, and SnakeSCEIAD, which uses the optimal parameters identified in SnakePOP to run the cell type machine learning and differential gene testing that is incorporated into the scEiaD sqlite database for the app. The code base for the web app (along with local installation instructions) is available at github.com/davemcg/scEiaD/plaeApp. Data is available for download at plae.nei.nih.gov (click "Data") and the Seurat object and metadata for the resource has been deposited at Zenodo with accession 10.5281/zenodo.7071682.
